I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-17 Paper/proofs.tex
update

	\begin{lemma}\label{lem:key1}

		Let $\vx^{*}$ be the input vector that for every node $a_i$, if $S(a_i)=1$, then $x_{a_i}$ is the maximal value of $a_i$, and if $S(a_i)=-1$, then $x_{a_i}$ is the minimal value of $a_i$ ($S(a_i)=0$, then $x_{a_i}$ be an arbitrary fixed value in the input interval of $a_i$).

		Then the input $\vx^*$ maximize all nodes $n$ that $S(n)=1$ and minimize all nodes $n$ that $S(n)=1$.
	\end{lemma}

	\begin{proof}
		We prove this lemma by induction. For input layer, this is true by definition. Suppose this lemma is true up to layer $L_i$, we need to show this for layer $L_{i+1}$. Let $c$ be a node in layer $L_{i+1}$, then by definition, we have the following equation: \begin{align*}
			c = \boldsymbol{bias}_c+\sum_{b\in L}W_{bc}\ReLU(b).
		\end{align*}

		Therefore we have the following inequalities:\begin{align*}
				\max(\Val(c))\leq\sum_{b \in L_i, W_{bc n}>0}W_{bc} \max(\Val(\hat{b})) + \sum_{b \in L_i, W_{bc}<0}W_{bc} \min(\Val(\hat{b})) + \boldsymbol{bisa}_c\\
			\min(\Val(c))\geq\sum_{b \in L_i, W_{bc n}>0}W_{bc} \min(\Val(\hat{b})) + \sum_{b \in L_i, W_{bc}<0}W_{bc} \max(\Val(\hat{b})) + \boldsymbol{bisa}_c
		\end{align*}

		Now, if $S(c)=1$, we consider different case of $b$. If $S(b)=1$, then by Lemma \ref{lem:sign}, $W_{bc}>0$; if $S(b)=-1$, then $W{bc}<0$; if $S(b)=0$, then $W_{bc}=0$.

		Combining with

	\begin{lemma}
		\label{lemma2}
		Let $L,L'$ be consecutive layers of a compensation free DNN, and $n \in L'$. 
		Then:
		$$ S(n) \max(S(n) n)=\sum_{m \in L}W_{m n} S(m) \max(S(m) \hat{m})+bias_n \text{ and }$$
		$$S(n) \max(S(n)\hat{n}_s)=\ReLU(\sum_{m \in L}W_{m n} S(n)\max(S(n)\hat{m}_s)+bias_n)$$

		Similarly for minimal value.

		%	for minimal value,	
		%	$ \min(n)=\sum_{m \in L}W_{m n)\min(\)+bias_n $ and
			%	$ \min(\hat{n}_s)=ReLU(\sum_{m \in L}W_{m n)\min(\hat{m}s}+bias_n)$
		\end{lemma}

		\begin{proof}
			We choose $\boldsymbol{x}^*$ maximizing each $a_s \in \{a,-a\}$ in the input layer $L=L_0$,
			that is setting the value for node $a$ be $\max(a)$ for $S(a)=1$ and $\min(a)$ for $S(a)=-1$.
			At configuration $\boldsymbol{x}^*$, every $a_s(\boldsymbol{x}^*))=\max_{\boldsymbol{x}} {a_s(\boldsymbol{x}))}=\max(a_s)$ is maximized.

			Consider any $b$ in the first hidden layer $L_1$.
			Assume first that $S(b)=+1$.
			We have $\Val(b_s)= \Val(b) = \sum_{a \in L_0} W_{ab} \Val(a) + bias_b$.
			If every $W_{ab} \Val(a)$ is maximized in the same time, 
			then $\Val(b_s)= \Val(b)$ is also maximized.

			We claim that this is the case at 
			configuration $\boldsymbol{x}^*$.
			Indeed, if $W_{ab}=0$, any $\Val(a)$ can be chosen to maximize 
			$W_{ab}\Val(a)$. If $W_{ab}>0$ then $S(a) = +1$ by Lemma \ref{lemma1}, since
			$S(b)=+1$. We have $\max(a_s) = \max(a)$, maximizing $W_{ab}\Val(a)$.
			Finally, if $W_{a b}<0$ then $S(a_s) = -1$, and we have 
			$\max(a_s) = min(a)$, maximizing $W_{ab}\Val(a)$.

			Hence $b$ reaches is maximum in configuration $\boldsymbol{x}^*$, and the value satisfies: 
			$$\max(b)=\sum_{m \in L}W_{a b} \max(a_s)+bias_b$$

			The case $S(b)=-1$ is symmetric:
			$\max(b_s)= -\min(b)$ is reached when minimizing $b$, which is also 
			satisfied at $\boldsymbol{x}^*$.
			In both case, $\Val(b_s)$ is maximized at $\boldsymbol{x}^*$. Notice that $\boldsymbol{x}^*$ does not depend upon $b$, so choosing $\boldsymbol{x}^*$ uniformly maximize all $\Val(b_s)$.

			This implies that $\ReLU(b_s)$ is maximized at $\boldsymbol{x}^*$, and its value is 
			$\ReLU(\sum_{m \in L}W_{a }\max(a_s)+bias_b)$, again uniformly reached over all $b$ at configuration $\boldsymbol{x}^*$.

			We proceed by induction over every layer of the DNN till the output layer, proving that 
			$\max(z_s)=\max(z)$ is reached at configuration $\boldsymbol{x}^*$.

			The case for $\min$ is similar, with another input vector $\boldsymbol{x}^\#$ instead of $\boldsymbol{x}^*$.
			\hfill $\square$
		\end{proof}

		As direct corollary of Lemma \ref{lemma2}, we obtain:

@@

		Notice that for all $x$, either $S(x)=+1$ and 
		$(\min(x),\max(x))=(\min(x_s),\max(x_s))$, 
		or $S(x)=-1$ and $(\min(x),\max(x))=(-\max(x_s),-\min(x_s))$.
		In both case, these two bounds are found in Lemma \ref{lemma2},
		from configuration $\boldsymbol{x}^*$ and $\boldsymbol{x}^\#$.

@@

				We define $\boldsymbol{x}_I^*$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=-1$). Similarly, we define $\boldsymbol{x}_I^\#$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=-1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=1$).

@@

					If $S(b)=-1$, then the similar results hold, replacing $\boldsymbol{x}^*_I$ by $\boldsymbol{x}^\#_I$. 
				\end{lemma}

@@

					1. for any fixed $\boldsymbol{x}^0_J$ and for $\boldsymbol{x}^*_I$, the value for $c$ in the MILP model is a fixed number, and naturally $\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)=\LB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)$. The same is true for $\boldsymbol{x}^\#_I$.

@@

					2. for any fixed $\boldsymbol{x}^0_J$, if $S(c)=1$,  $\max_{\boldsymbol{x}_I} \UB_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)=\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)= \Val_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c) = \max_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)$. The similar for $S(c)=-1$, $\langle \LB,\min\rangle$, $x^\#_I$: we can change two of them at the same time.

>>>>

	\begin{definition}[Star and Sharp]
	\begin{enumerate}
		\item  $\vx^{*}$ is a input vector that for every node $a_i$, if $S(a_i)=1$, then $x_{a_i}$ is the maximal value of $a_i$, and if $S(a_i)=-1$, then $x_{a_i}$ is the minimal value of $a_i$ (if $S(a_i)=0$, then $x_{a_i}$ be an arbitrary fixed value in the input interval of $a_i$).

		\item  $\vx^{\sharp}$ is a input vector that for every node $a_i$, if $S(a_i)=-1$, then $x_{a_i}$ is the maximal value of $a_i$, and if $S(a_i)=1$, then $x_{a_i}$ is the minimal value of $a_i$ (if $S(a_i)=0$, then $x_{a_i}$ be an arbitrary fixed value in the input interval of $a_i$).
	\end{enumerate}

	\end{definition}

	\begin{lemma}\label{lem:key1}\label{lemma2}

		The input $\vx^*$ maximize all nodes $n$ that $S(n)=1$ and minimize all nodes $n$ that $S(n)=-1$.

		Similarly, the input $\vx^\sharp$ maximize all nodes $n$ that $S(n)=-1$ and minimize all nodes $n$ that $S(n)=1$.
	\end{lemma}

	\begin{proof}
		We prove this lemma by induction on layers. For input layer, this is true by definition. Suppose this lemma is true up to layer $L_i$, we need to show this for layer $L_{i+1}$. Let $c$ be a node in layer $L_{i+1}$, then by definition, we have the following equation: \begin{align*}
			c = \boldsymbol{bias}_c+\sum_{b\in L}W_{bc}\ReLU(b).
		\end{align*}

		Therefore we have the following inequalities:\begin{align*}
				\max(c)\leq\sum_{b \in L_i, W_{bc }>0}W_{bc} \max(\hat{b}) + \sum_{b \in L_i, W_{bc}<0}W_{bc} \min(\hat{b}) + \boldsymbol{bisa}_c\\
			\min(c)\geq\sum_{b \in L_i, W_{bc}>0}W_{bc} \min(\hat{b}) + \sum_{b \in L_i, W_{bc}<0}W_{bc} \max(\hat{b}) + \boldsymbol{bisa}_c
		\end{align*}

		Now, if $S(c)=1$, we consider different case of $b$. If $S(b)=1$, then by Lemma \ref{lem:sign}, $W_{bc}>0$; if $S(b)=-1$, then $W_{bc}<0$; if $S(b)=0$, then $W_{bc}=0$.

		By induction hypothesis, when input vector is $\vx^{*}$, for every $b$, if $S(b)=1$, then $\Val_{\vx}(b)=\max(b)$ and if $S(b)=-1$, then $\Val_{\vx}(b)=\min(b)$.

		So combining formulas and above two facts, we will have that when the input vector is $\vx^*$, \begin{align*}
			\max(c)&\geq\Val_{\vx^*}(c) \\
			&= \sum_{b \in L_i, W_{bc}>0}W_{bc} \max(\hat{b}) + \sum_{b \in L_i, W_{bc}<0}W_{bc} \min(\hat{b}) + \boldsymbol{bisa}_c\\
			&\geq\max(c)
		\end{align*}  Hence $\max(c)=\Val_{\vx^*}(c)$.

		Similarly, we can prove if $S(c)=-1$, $\min(c)=\Val_{\vx^*}(c)$. And similarly, we can prove this for $x^\sharp$. This is the induction step. 
	\end{proof}

%	\begin{lemma}
%		\label{lemma2}
%		Let $L,L'$ be consecutive layers of a compensation free DNN, and $n \in L'$. 
%		Then:
%		$$ S(n) \max(S(n) n)=\sum_{m \in L}W_{m n} S(m) \max(S(m) \hat{m})+bias_n \text{ and }$$
%		$$S(n) \max(S(n)\hat{n}_s)=\ReLU(\sum_{m \in L}W_{m n} S(n)\max(S(n)\hat{m}_s)+bias_n)$$
%		
%		Similarly for minimal value.
%		
%		%	for minimal value,	
%		%	$ \min(n)=\sum_{m \in L}W_{m n)\min(\)+bias_n $ and
%			%	$ \min(\hat{n}_s)=ReLU(\sum_{m \in L}W_{m n)\min(\hat{m}s}+bias_n)$
%		\end{lemma}
%		
%		\begin{proof}
%			We choose $\boldsymbol{x}^*$ maximizing each $a_s \in \{a,-a\}$ in the input layer $L=L_0$,
%			that is setting the value for node $a$ be $\max(a)$ for $S(a)=1$ and $\min(a)$ for $S(a)=-1$.
%			At configuration $\boldsymbol{x}^*$, every $a_s(\boldsymbol{x}^*))=\max_{\boldsymbol{x}} {a_s(\boldsymbol{x}))}=\max(a_s)$ is maximized.  
%			
%			Consider any $b$ in the first hidden layer $L_1$.
%			Assume first that $S(b)=+1$.
%			We have $\Val(b_s)= \Val(b) = \sum_{a \in L_0} W_{ab} \Val(a) + bias_b$.
%			If every $W_{ab} \Val(a)$ is maximized in the same time, 
%			then $\Val(b_s)= \Val(b)$ is also maximized. 
%			
%			We claim that this is the case at 
%			configuration $\boldsymbol{x}^*$.
%			Indeed, if $W_{ab}=0$, any $\Val(a)$ can be chosen to maximize 
%			$W_{ab}\Val(a)$. If $W_{ab}>0$ then $S(a) = +1$ by Lemma \ref{lemma1}, since
%			$S(b)=+1$. We have $\max(a_s) = \max(a)$, maximizing $W_{ab}\Val(a)$.
%			Finally, if $W_{a b}<0$ then $S(a_s) = -1$, and we have 
%			$\max(a_s) = min(a)$, maximizing $W_{ab}\Val(a)$.
%			
%			Hence $b$ reaches is maximum in configuration $\boldsymbol{x}^*$, and the value satisfies: 
%			$$\max(b)=\sum_{m \in L}W_{a b} \max(a_s)+bias_b$$
%			
%			The case $S(b)=-1$ is symmetric:
%			$\max(b_s)= -\min(b)$ is reached when minimizing $b$, which is also 
%			satisfied at $\boldsymbol{x}^*$.
%			In both case, $\Val(b_s)$ is maximized at $\boldsymbol{x}^*$. Notice that $\boldsymbol{x}^*$ does not depend upon $b$, so choosing $\boldsymbol{x}^*$ uniformly maximize all $\Val(b_s)$.
%			
%			This implies that $\ReLU(b_s)$ is maximized at $\boldsymbol{x}^*$, and its value is 
%			$\ReLU(\sum_{m \in L}W_{a }\max(a_s)+bias_b)$, again uniformly reached over all $b$ at configuration $\boldsymbol{x}^*$.
%			
%			We proceed by induction over every layer of the DNN till the output layer, proving that 
%			$\max(z_s)=\max(z)$ is reached at configuration $\boldsymbol{x}^*$.
%			
%			The case for $\min$ is similar, with another input vector $\boldsymbol{x}^\sharp$ instead of $\boldsymbol{x}^*$.
%			\hfill $\square$
%		\end{proof}

		As direct corollary of Lemma \ref{lem:key1}, we obtain:

@@

		Notice that for all $x$, either $S(x)=+1$ and 
		$(\min(x),\max(x))=(\min(x_s),\max(x_s))$, 
		or $S(x)=-1$ and $(\min(x),\max(x))=(-\max(x_s),-\min(x_s))$.
		In both case, these two bounds are found in Lemma \ref{lemma2},
		from configuration $\boldsymbol{x}^*$ and $\boldsymbol{x}^\sharp$.

@@

				We define $\boldsymbol{x}_I^*$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=-1$). Similarly, we define $\boldsymbol{x}_I^\sharp$ to be the input vector such that for all $a_i \in I$, the value $x_{a_i}$ for $a_i$ gets its maximal value (if $S(a_i)=-1$ or $0$) or $a_i$ gets its minimal value (if $S(a_i)=1$).

@@

					If $S(b)=-1$, then the similar results hold, replacing $\boldsymbol{x}^*_I$ by $\boldsymbol{x}^\sharp_I$. 
				\end{lemma}

@@

					1. for any fixed $\boldsymbol{x}^0_J$ and for $\boldsymbol{x}^*_I$, the value for $c$ in the MILP model is a fixed number, and naturally $\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)=\LB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)$. The same is true for $\boldsymbol{x}^\sharp_I$.

@@

					2. for any fixed $\boldsymbol{x}^0_J$, if $S(c)=1$,  $\max_{\boldsymbol{x}_I} \UB_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)=\UB_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c)= \Val_{\boldsymbol{x}^*_I,\boldsymbol{x}^0_J}(c) = \max_{\boldsymbol{x}_I} \Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(c)$. The similar for $S(c)=-1$, $\langle \LB,\min\rangle$, $x^\sharp_I$: we can change two of them at the same time.
